---
title: "UCI Online Shopping Data Documentation"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
work_dir <- here()

# Table helpers for nicer PDF tables
library(kableExtra)

nice_var_table <- function(df,
                           caption = NULL,
                           var_width = "2.5cm",
                           desc_width = "10.5cm",
                           wrap_width = 40) {
  # Apply explicit line breaks to the description column (2nd column)
  if (ncol(df) >= 2) {
    df[[2]] <- kableExtra::linebreak(df[[2]], align = "l")
  }

  knitr::kable(
    df,
    booktabs = TRUE,
    linesep = "\\hline",
    caption = caption,
    longtable = TRUE,
    escape = FALSE
  ) |>
    kableExtra::kable_styling(
      full_width = FALSE,
      font_size = 8,
      latex_options = c("repeat_header"),
      position = "center"
    ) |>
    kableExtra::column_spec(1, width = var_width) |>
    kableExtra::column_spec(2, width = desc_width)
}
```

# Overview

This document describes the **UCI Online Shoppers Purchasing Intention Dataset** and the derived datasets stored in the `shopping/data` subfolders.  
The original data come from the UCI Machine Learning Repository (Online Shoppers Purchasing Intention Data Set; see UCI entry and related paper, e.g. via Semantic Scholar).

The following datasets are covered here:

- **Raw UCI data**: `data/raw_data/uci_shopping.csv`
- **Preprocessed UCI data**: `data/intermediate_data/01_uci_preprocessed.csv`
- **Uniformly censored UCI data**: `data/intermediate_data/02_uci_uniform_censor.csv`
- **Randomly censored UCI data**: `data/intermediate_data/03_uci_random_censor.csv`

The code that creates these datasets lives in the `shopping/code` folder:

- `00_uci_EDA.Rmd`
- `01_uci_data_preprocessing.R`
- `02_uci_censor_uniform_time.R`
- `03_uci_censor_random_time.R`

Below, we first describe the variables in the **raw** UCI dataset and then explain how each intermediate dataset is constructed, including the key code used.

---

## 1. Raw UCI dataset: `uci_shopping.csv`

**File path**: `shopping/data/raw_data/uci_shopping.csv`  
**Source**: UCI Machine Learning Repository, Online Shoppers Purchasing Intention Data Set (ID 468).

```{r, message = FALSE}
uci_raw <- read_csv(file.path(work_dir, "shopping", "data", "raw_data", "uci_shopping.csv"),
                    show_col_types = FALSE)
glimpse(uci_raw)
```

### Variable descriptions 

The variables below follow the UCI dataset documentation and the explanations in `00_uci_EDA.Rmd`.

```{r raw_uci_variable_tables, echo = FALSE}
library(dplyr)

raw_continuous <- tibble::tribble(
  ~Variable, ~Description,
  "Administrative",          "Number of pages visited that are related to account management (e.g., account settings, order history).",
  "Administrative_Duration", "Total time (in seconds) spent on account management–related pages during the session.",
  "Informational",           "Number of informational pages visited (e.g., pages about the company, contact details, FAQs).",
  "Informational_Duration",  "Total time (in seconds) spent on informational pages.",
  "ProductRelated",          "Number of product-related pages visited (e.g., product listings, product detail pages, cart pages).",
  "ProductRelated_Duration", "Total time (in seconds) spent on product-related pages.",
  "BounceRates",             "Average bounce rate of the visited pages in the session; roughly 1 - Engagement Rate, where engagement is based on session length, key events, or multiple page views.",
  "ExitRates",               "Average exit rate of the visited pages; approximates (Number of Exits / Number of Page Views) x 100 aggregated over pages in the session.",
  "PageValues",              "Average page value of the visited pages; combines total revenue and goal value per unique page view according to Google Analytics definitions.",
  "SpecialDay",              "Continuous measure of closeness of the visit date to a special day (e.g., Valentine’s Day, Mother’s Day); higher values indicate visits closer to the special day."
)

raw_categorical <- tibble::tribble(
  ~Variable, ~Description,
  "Month",            "Month of the visit (e.g., \"Feb\", \"Mar\", \"May\").",
  "OperatingSystems", "Encoded identifier of the visitor’s operating system.",
  "Browser",          "Encoded identifier of the visitor’s web browser.",
  "Region",           "Encoded geographic region from which the session was started.",
  "TrafficType",      "Encoded traffic source indicating how the visitor arrived at the site (e.g., direct, referral, ad campaign).",
  "VisitorType",      "Visitor type; typically \"New_Visitor\", \"Returning_Visitor\", or \"Other\".",
  "Weekend",          "Logical indicator; TRUE if the session took place on a weekend, FALSE otherwise.",
  "Revenue",          "Logical indicator; TRUE if the session ended with a transaction (purchase), FALSE otherwise."
)

nice_var_table(raw_continuous, caption = "Raw UCI continuous variables",
               var_width = "4cm",
               desc_width = "12cm")

nice_var_table(raw_categorical, caption = "Raw UCI categorical variables")
```

---

## 2. Preprocessed UCI dataset: `01_uci_preprocessed.csv`

**File path**: `shopping/data/intermediate_data/01_uci_preprocessed.csv`  
**Generating code**: `shopping/code/01_uci_data_preprocessing.R`

### How `01_uci_preprocessed.csv` is created

Key steps from `01_uci_data_preprocessing.R`:

```{r, eval = FALSE}
library(tidyverse)
library(here)

work_dir <- here()
raw_data <- read_csv(file.path(work_dir, "shopping", "data", "raw_data", "uci_shopping.csv"))

# Summing up durations
uci_data_summed <- raw_data %>%
  mutate(TotalDuration = Administrative_Duration +
                          Informational_Duration +
                          ProductRelated_Duration) %>%
  select(-c(Administrative_Duration,
            Informational_Duration,
            ProductRelated_Duration)) %>%
  filter(TotalDuration != 0)  # filter out sessions with zero total duration

write.csv(uci_data_summed,
          file.path(work_dir, "shopping", "data", "intermediate_data",
                    "01_uci_preprocessed.csv"),
          row.names = FALSE)
```

### Variable changes in `01_uci_preprocessed.csv`

```{r uci_preprocessed_variable_table, echo = FALSE}
preprocessed_changes <- tibble::tribble(
  ~Variable,   ~Description,
  "Administrative_Duration. (Removed)", "Original per-session time on administrative pages; removed after being summed into TotalDuration.",
  "Informational_Duration (Removed)", "Original per-session time on informational pages; removed after being summed into TotalDuration.",
  "ProductRelated_Duration (Removed)", "Original per-session time on product-related pages; removed after being summed into TotalDuration.",
  "TotalDuration (New)","Sum of Administrative_Duration + Informational_Duration + ProductRelated_Duration; strictly positive in this dataset."
)

nice_var_table(
  preprocessed_changes,
  caption = "Changes from raw UCI data to the preprocessed dataset",
  var_width = "4cm",
  desc_width = "10cm"
)
```

```{r, message = FALSE}
uci_pre <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                              "01_uci_preprocessed.csv"),
                    show_col_types = FALSE)
glimpse(uci_pre)
```

---

## 3. Uniformly censored UCI dataset: `02_uci_uniform_censor.csv`

**File path**: `shopping/data/intermediate_data/02_uci_uniform_censor.csv`  
**Generating code**: `shopping/code/02_uci_censor_uniform_time.R`

### How `02_uci_uniform_censor.csv` is created

Key steps from `02_uci_censor_uniform_time.R`:

```{r, eval = FALSE}
library(tidyverse)
library(here)

work_dir <- here()
uci_data <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                               "01_uci_preprocessed.csv"))

# Introducing censorship with a single uniform time threshold
set.seed(123)
time_threshold <- runif(1, min = 10, max = 1800)  # between 10 seconds and 30 minutes

uci_transformed <- uci_data %>%
  mutate(
    TimeThreshold = time_threshold,
    CensorTime   = ifelse(TotalDuration >= time_threshold,
                          time_threshold,
                          TotalDuration),
    Censor       = ifelse(TotalDuration >= time_threshold, 1, 0)
  )

write.csv(uci_transformed,
          file.path(work_dir, "shopping", "data", "intermediate_data",
                    "02_uci_uniform_censor.csv"),
          row.names = FALSE)
```


### New variables in `02_uci_uniform_censor.csv`

Starting from `01_uci_preprocessed.csv`, this dataset adds three new variables:

- **`TimeThreshold`**:
  - A scalar threshold (same value for all observations).
  - Drawn from a continuous uniform distribution on \([10, 1800]\) seconds.
  - Represents a global censoring time.

- **`CensorTime`**:
  - The observed time after applying censoring:
    \[
    \text{CensorTime} =
      \begin{cases}
        \text{TotalDuration}, & \text{if } \text{TotalDuration} < \text{TimeThreshold}, \\
        \text{TimeThreshold}, & \text{if } \text{TotalDuration} \ge \text{TimeThreshold}.
      \end{cases}
    \]

- **`Censor`**:
  - Indicator of whether an observation is censored at `TimeThreshold`:
    - `1` if `TotalDuration >= TimeThreshold` (censored).
    - `0` otherwise (fully observed).

```{r uci_uniform_censor_variable_table, echo = FALSE}
uniform_censor_vars <- tibble::tribble(
  ~Variable,       ~Description,
  "TimeThreshold", "Scalar censoring threshold (same value for all observations), drawn from a continuous uniform distribution on [10, 1800] seconds; represents a global censoring time.",
  "CensorTime",    "Observed time after applying censoring: equals TotalDuration if TotalDuration < TimeThreshold, otherwise equals TimeThreshold.",
  "Censor",        "Censoring indicator: 1 if TotalDuration >= TimeThreshold (censored at the threshold), 0 otherwise."
)

nice_var_table(
  uniform_censor_vars,
  caption = "Additional variables in the uniformly censored UCI dataset",
  var_width = "4cm",
  desc_width = "11cm"
)
```

All other variables from `01_uci_preprocessed.csv` are preserved unchanged.

```{r, message = FALSE}
uci_uniform <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                                  "02_uci_uniform_censor.csv"),
                        show_col_types = FALSE)
glimpse(uci_uniform)
```

---

## 4. Randomly censored UCI dataset: `03_uci_random_censor.csv`

**File path**: `shopping/data/intermediate_data/03_uci_random_censor.csv`  
**Generating code**: `shopping/code/03_uci_censor_random_time.R`

### How `03_uci_random_censor.csv` is created

Key steps from `03_uci_censor_random_time.R`:

```{r, eval = FALSE}
library(tidyverse)
library(here)

work_dir <- here()
uci_data <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                               "01_uci_preprocessed.csv"))

# Introducing censorship with row-specific random thresholds
set.seed(123)
time_threshold <- runif(nrow(uci_data), min = 10, max = 1800)  # one threshold per row

uci_transformed <- uci_data %>%
  mutate(
    TimeThreshold = time_threshold,
    CensorTime   = ifelse(TotalDuration >= time_threshold,
                          time_threshold,
                          TotalDuration),
    Censor       = ifelse(TotalDuration >= time_threshold, 1, 0)
  )

write.csv(uci_transformed,
          file.path(work_dir, "shopping", "data", "intermediate_data",
                    "03_uci_random_censor.csv"),
          row.names = FALSE)
```

### New variables in `03_uci_random_censor.csv`

Relative to `01_uci_preprocessed.csv`, the same three censoring variables are added as in the uniform case, but with **observation-specific** thresholds:

- **`TimeThreshold`**:
  - A numeric vector of length \(n\), with one value per observation.
  - Each value is drawn independently from \([10, 1800]\) seconds.

- **`CensorTime`**:
  - Row-wise censored time:
    \[
    \text{CensorTime}_i =
      \begin{cases}
        \text{TotalDuration}_i, & \text{if } \text{TotalDuration}_i < \text{TimeThreshold}_i, \\
        \text{TimeThreshold}_i, & \text{if } \text{TotalDuration}_i \ge \text{TimeThreshold}_i.
      \end{cases}
    \]

- **`Censor`**:
  - Row-wise censoring indicator:
    - `1` if `TotalDuration_i >= TimeThreshold_i` (censored).
    - `0` otherwise.

```{r uci_random_censor_variable_table, echo = FALSE}
random_censor_vars <- tibble::tribble(
  ~Variable,       ~Description,
  "TimeThreshold", "Numeric vector of length n (one value per observation); each value is drawn independently from a uniform distribution on [10, 1800] seconds.",
  "CensorTime",    "Row-wise censored time: equals TotalDuration_i if TotalDuration_i < TimeThreshold_i, otherwise equals TimeThreshold_i.",
  "Censor",        "Row-wise censoring indicator: 1 if TotalDuration_i >= TimeThreshold_i (censored), 0 otherwise."
)

nice_var_table(
  random_censor_vars,
  caption = "Additional variables in the randomly censored UCI dataset",
  var_width = "4cm",
  desc_width = "11cm"
)
```

All other variables (from `01_uci_preprocessed.csv`) remain unchanged.

```{r, message = FALSE}
uci_random <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                                 "03_uci_random_censor.csv"),
                       show_col_types = FALSE)
glimpse(uci_random)
```

---

## 5. Relationship between datasets and code files

- **Exploratory data analysis and variable explanations** for the raw UCI dataset are in `shopping/code/00_uci_EDA.Rmd`.  
- **Creation of `01_uci_preprocessed.csv`** is handled by `shopping/code/01_uci_data_preprocessing.R`, which:
  - Reads `uci_shopping.csv`.
  - Sums duration variables into `TotalDuration`.
  - Drops zero–total-duration sessions.
  - Writes the result to `01_uci_preprocessed.csv`.
- **Creation of `02_uci_uniform_censor.csv`** is handled by `shopping/code/02_uci_censor_uniform_time.R`, which:
  - Reads `01_uci_preprocessed.csv`.
  - Draws a **single** `TimeThreshold` from \([10, 1800]\) seconds.
  - Creates `CensorTime` and `Censor` based on that threshold.
  - Writes the result to `02_uci_uniform_censor.csv`.
- **Creation of `03_uci_random_censor.csv`** is handled by `shopping/code/03_uci_censor_random_time.R`, which:
  - Reads `01_uci_preprocessed.csv`.
  - Draws **row-specific** `TimeThreshold` values from \([10, 1800]\) seconds.
  - Creates `CensorTime` and `Censor` for each row.
  - Writes the result to `03_uci_random_censor.csv`.

Together, these files document the full pipeline from the original UCI Online Shoppers Purchasing Intention data to the censored datasets used for subsequent analysis.

