---
title: "UCI Online Shopping Data Documentation"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
work_dir <- here()

# Table helpers for nicer PDF tables
library(kableExtra)

nice_var_table <- function(df,
                           caption = NULL,
                           var_width = "2.5cm",
                           desc_width = "10.5cm",
                           wrap_width = 40) {
  # Apply explicit line breaks to the description column (2nd column)
  if (ncol(df) >= 2) {
    df[[2]] <- kableExtra::linebreak(df[[2]], align = "l")
  }

  knitr::kable(
    df,
    booktabs = TRUE,
    linesep = "\\hline",
    caption = caption,
    longtable = TRUE,
    escape = FALSE
  ) |>
    kableExtra::kable_styling(
      full_width = FALSE,
      font_size = 8,
      latex_options = c("repeat_header"),
      position = "center"
    ) |>
    kableExtra::column_spec(1, width = var_width) |>
    kableExtra::column_spec(2, width = desc_width)
}
```

# Overview

This document describes the **UCI Online Shoppers Purchasing Intention Dataset** and the derived datasets stored in the `shopping/data` subfolders.  
The original data come from the UCI Machine Learning Repository (Online Shoppers Purchasing Intention Data Set; see UCI entry and related paper, e.g. via Semantic Scholar).

The following datasets are covered here:

- **Final data (analysis-ready)**: `data/final_data/shopping_data_final.csv`
- **Raw UCI data**: `data/raw_data/uci_shopping.csv`
- **Preprocessed UCI data**: `data/intermediate_data/01_uci_preprocessed.csv`
- **Randomly censored UCI data**: `data/intermediate_data/02_uci_random_censor.csv`

The code that creates these datasets lives in the `shopping/code` folder:

- `00_uci_EDA.Rmd`
- `01_uci_data_preprocessing.R`
- `02_uci_censor_random_time.R`

Below, we describe the **final** dataset and the **relationship between datasets and code files**; detailed descriptions of the individual datasets (raw, preprocessed, and randomly censored) are in the **Appendix**.

---

## 1. Final dataset: `shopping_data_final.csv`

**File path**: `shopping/data/final_data/shopping_data_final.csv`  
**Generating code**: `shopping/code/02_uci_censor_random_time.R`

This is the **analysis-ready** dataset produced by the pipeline. It has the same structure and content as the randomly censored intermediate file `02_uci_random_censor.csv`: preprocessed UCI sessions with row-specific random censoring applied to `TotalDuration`, plus the censoring variables `TimeThreshold`, `CensorTime`, and `Censor`. The final file is written to `data/final_data/` for use in downstream analysis and reporting.

### Variable descriptions

The tables below describe every variable in the final dataset. Continuous and count variables are listed first; categorical variables follow.

```{r final_data_variable_tables, echo = FALSE}
final_continuous <- tibble::tribble(
  ~Variable,       ~Description,
  "Administrative",   "Number of pages visited that are related to account management (e.g., account settings, order history).",
  "Informational",    "Number of informational pages visited (e.g., pages about the company, contact details, FAQs).",
  "ProductRelated",   "Number of product-related pages visited (e.g., product listings, product detail pages, cart pages).",
  "BounceRates",      "Average bounce rate of the visited pages in the session; roughly 1 - Engagement Rate.",
  "ExitRates",        "Average exit rate of the visited pages; approximates (Number of Exits / Number of Page Views) x 100 over pages in the session.",
  "PageValues",       "Average page value of the visited pages; combines total revenue and goal value per unique page view (Google Analytics definitions).",
  "SpecialDay",       "Continuous measure of closeness of the visit date to a special day (e.g., Valentine's Day, Mother's Day); higher values indicate visits closer to the special day.",
  "TotalDuration",    "Total time (in seconds) spent in the session (sum of administrative, informational, and product-related page durations); strictly positive.",
  "TimeThreshold",    "Row-specific censoring threshold (seconds); each value drawn independently from a uniform distribution on [10, 1800].",
  "CensorTime",       "Observed time (seconds) after censoring: equals TotalDuration if TotalDuration < TimeThreshold, otherwise equals TimeThreshold.",
  "Censor",           "Censoring indicator: 1 if TotalDuration >= TimeThreshold (censored), 0 otherwise (fully observed)."
)

final_categorical <- tibble::tribble(
  ~Variable,         ~Description,
  "Month",            "Month of the visit (e.g., \"Feb\", \"Mar\", \"May\").",
  "OperatingSystems", "Encoded identifier of the visitor's operating system.",
  "Browser",          "Encoded identifier of the visitor's web browser.",
  "Region",           "Encoded geographic region from which the session was started.",
  "TrafficType",      "Encoded traffic source indicating how the visitor arrived at the site (e.g., direct, referral, ad campaign).",
  "VisitorType",      "Visitor type; typically \"New_Visitor\", \"Returning_Visitor\", or \"Other\".",
  "Weekend",          "Logical; TRUE if the session took place on a weekend, FALSE otherwise.",
  "Revenue",          "Logical; TRUE if the session ended with a transaction (purchase), FALSE otherwise."
)

nice_var_table(final_continuous, caption = "Final dataset: continuous and count variables",
               var_width = "3.2cm", desc_width = "11.5cm")
nice_var_table(final_categorical, caption = "Final dataset: categorical variables",
               var_width = "3.2cm", desc_width = "11.5cm")
```

```{r final_data_glimpse, message = FALSE}
shopping_final <- read_csv(file.path(work_dir, "shopping", "data", "final_data", 
                                     "shopping_data_final.csv"),
                           show_col_types = FALSE)
glimpse(shopping_final)
```

---

# Appendix

Detailed descriptions of each dataset in the pipeline are given below.

## A. Relationship between datasets and code files

- **Exploratory data analysis and variable explanations** for the raw UCI dataset are in `shopping/code/00_uci_EDA.Rmd`.  
- **Creation of `01_uci_preprocessed.csv`** is handled by `shopping/code/01_uci_data_preprocessing.R`, which:
  - Reads `uci_shopping.csv`.
  - Sums duration variables into `TotalDuration`.
  - Drops zero–total-duration sessions.
  - Writes the result to `01_uci_preprocessed.csv`.
- **Creation of `02_uci_random_censor.csv` and `shopping_data_final.csv`** is handled by `shopping/code/02_uci_censor_random_time.R`, which:
  - Reads `01_uci_preprocessed.csv`.
  - Draws **row-specific** `TimeThreshold` values from \([10, 1800]\) seconds.
  - Creates `CensorTime` and `Censor` for each row.
  - Writes the result to `02_uci_random_censor.csv` (intermediate) and to 
  \
  `shopping/data/final_data/shopping_data_final.csv` (final, analysis-ready).

Together, these files document the full pipeline from the original UCI Online Shoppers Purchasing Intention data to the final dataset used for subsequent analysis.

## B. Raw UCI dataset: `uci_shopping.csv`

**File path**: `shopping/data/raw_data/uci_shopping.csv`  
**Source**: UCI Machine Learning Repository, Online Shoppers Purchasing Intention Data Set (ID 468).

```{r, message = FALSE}
uci_raw <- read_csv(file.path(work_dir, "shopping", "data", "raw_data", "uci_shopping.csv"),
                    show_col_types = FALSE)
glimpse(uci_raw)
```

### Variable descriptions 

The variables below follow the UCI dataset documentation and the explanations in `00_uci_EDA.Rmd`.

```{r raw_uci_variable_tables, echo = FALSE}
library(dplyr)

raw_continuous <- tibble::tribble(
  ~Variable, ~Description,
  "Administrative",          "Number of pages visited that are related to account management (e.g., account settings, order history).",
  "Administrative_Duration", "Total time (in seconds) spent on account management–related pages during the session.",
  "Informational",           "Number of informational pages visited (e.g., pages about the company, contact details, FAQs).",
  "Informational_Duration",  "Total time (in seconds) spent on informational pages.",
  "ProductRelated",          "Number of product-related pages visited (e.g., product listings, product detail pages, cart pages).",
  "ProductRelated_Duration", "Total time (in seconds) spent on product-related pages.",
  "BounceRates",             "Average bounce rate of the visited pages in the session; roughly 1 - Engagement Rate, where engagement is based on session length, key events, or multiple page views.",
  "ExitRates",               "Average exit rate of the visited pages; approximates (Number of Exits / Number of Page Views) x 100 aggregated over pages in the session.",
  "PageValues",              "Average page value of the visited pages; combines total revenue and goal value per unique page view according to Google Analytics definitions.",
  "SpecialDay",              "Continuous measure of closeness of the visit date to a special day (e.g., Valentine’s Day, Mother’s Day); higher values indicate visits closer to the special day."
)

raw_categorical <- tibble::tribble(
  ~Variable, ~Description,
  "Month",            "Month of the visit (e.g., \"Feb\", \"Mar\", \"May\").",
  "OperatingSystems", "Encoded identifier of the visitor’s operating system.",
  "Browser",          "Encoded identifier of the visitor’s web browser.",
  "Region",           "Encoded geographic region from which the session was started.",
  "TrafficType",      "Encoded traffic source indicating how the visitor arrived at the site (e.g., direct, referral, ad campaign).",
  "VisitorType",      "Visitor type; typically \"New_Visitor\", \"Returning_Visitor\", or \"Other\".",
  "Weekend",          "Logical indicator; TRUE if the session took place on a weekend, FALSE otherwise.",
  "Revenue",          "Logical indicator; TRUE if the session ended with a transaction (purchase), FALSE otherwise."
)

nice_var_table(raw_continuous, caption = "Raw UCI continuous variables",
               var_width = "4cm",
               desc_width = "12cm")

nice_var_table(raw_categorical, caption = "Raw UCI categorical variables")
```

---

## C. Preprocessed UCI dataset: `01_uci_preprocessed.csv`

**File path**: `shopping/data/intermediate_data/01_uci_preprocessed.csv`  
**Generating code**: `shopping/code/01_uci_data_preprocessing.R`

### How `01_uci_preprocessed.csv` is created

Key steps from `01_uci_data_preprocessing.R`:

```{r, eval = FALSE}
library(tidyverse)
library(here)

work_dir <- here()
raw_data <- read_csv(file.path(work_dir, "shopping", "data", "raw_data", "uci_shopping.csv"))

# Summing up durations
uci_data_summed <- raw_data %>%
  mutate(TotalDuration = Administrative_Duration +
                          Informational_Duration +
                          ProductRelated_Duration) %>%
  select(-c(Administrative_Duration,
            Informational_Duration,
            ProductRelated_Duration)) %>%
  filter(TotalDuration != 0)  # filter out sessions with zero total duration

write.csv(uci_data_summed,
          file.path(work_dir, "shopping", "data", "intermediate_data",
                    "01_uci_preprocessed.csv"),
          row.names = FALSE)
```

### Variable changes in `01_uci_preprocessed.csv`

```{r uci_preprocessed_variable_table, echo = FALSE}
preprocessed_changes <- tibble::tribble(
  ~Variable,   ~Description,
  "Administrative_Duration. (Removed)", "Original per-session time on administrative pages; removed after being summed into TotalDuration.",
  "Informational_Duration (Removed)", "Original per-session time on informational pages; removed after being summed into TotalDuration.",
  "ProductRelated_Duration (Removed)", "Original per-session time on product-related pages; removed after being summed into TotalDuration.",
  "TotalDuration (New)","Sum of Administrative_Duration + Informational_Duration + ProductRelated_Duration; strictly positive in this dataset."
)

nice_var_table(
  preprocessed_changes,
  caption = "Changes from raw UCI data to the preprocessed dataset",
  var_width = "4cm",
  desc_width = "10cm"
)
```

```{r, message = FALSE}
uci_pre <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                              "01_uci_preprocessed.csv"),
                    show_col_types = FALSE)
glimpse(uci_pre)
```

---

## D. Randomly censored UCI dataset: `02_uci_random_censor.csv`

**File path**: `shopping/data/intermediate_data/02_uci_random_censor.csv`  
**Generating code**: `shopping/code/02_uci_censor_random_time.R`

The same script also writes the analysis-ready copy to `shopping/data/final_data/shopping_data_final.csv`.

### How `02_uci_random_censor.csv` is created

Key steps from `02_uci_censor_random_time.R`:

```{r, eval = FALSE}
library(tidyverse)
library(here)

work_dir <- here()
uci_data <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                               "01_uci_preprocessed.csv"))

# Introducing censorship with row-specific random thresholds
set.seed(123)
time_threshold <- runif(nrow(uci_data), min = 10, max = 1800)  # one threshold per row

uci_transformed <- uci_data %>%
  mutate(
    TimeThreshold = time_threshold,
    CensorTime   = ifelse(TotalDuration >= time_threshold,
                          time_threshold,
                          TotalDuration),
    Censor       = ifelse(TotalDuration >= time_threshold, 1, 0)
  )

write.csv(uci_transformed,
          file.path(work_dir, "shopping", "data", "intermediate_data",
                    "02_uci_random_censor.csv"),
          row.names = FALSE)
write.csv(uci_transformed,
          file.path(work_dir, "shopping", "data", "final_data",
                    "shopping_data_final.csv"),
          row.names = FALSE)
```

### New variables in `02_uci_random_censor.csv`

Relative to `01_uci_preprocessed.csv`, the same three censoring variables are added as in the uniform case, but with **observation-specific** thresholds:

- **`TimeThreshold`**:
  - A numeric vector of length \(n\), with one value per observation.
  - Each value is drawn independently from \([10, 1800]\) seconds.

- **`CensorTime`**:
  - Row-wise censored time:
    \[
    \text{CensorTime}_i =
      \begin{cases}
        \text{TotalDuration}_i, & \text{if } \text{TotalDuration}_i < \text{TimeThreshold}_i, \\
        \text{TimeThreshold}_i, & \text{if } \text{TotalDuration}_i \ge \text{TimeThreshold}_i.
      \end{cases}
    \]

- **`Censor`**:
  - Row-wise censoring indicator:
    - `1` if `TotalDuration_i >= TimeThreshold_i` (censored).
    - `0` otherwise.

```{r uci_random_censor_variable_table, echo = FALSE}
random_censor_vars <- tibble::tribble(
  ~Variable,       ~Description,
  "TimeThreshold", "Numeric vector of length n (one value per observation); each value is drawn independently from a uniform distribution on [10, 1800] seconds.",
  "CensorTime",    "Row-wise censored time: equals TotalDuration_i if TotalDuration_i < TimeThreshold_i, otherwise equals TimeThreshold_i.",
  "Censor",        "Row-wise censoring indicator: 1 if TotalDuration_i >= TimeThreshold_i (censored), 0 otherwise."
)

nice_var_table(
  random_censor_vars,
  caption = "Additional variables in the randomly censored UCI dataset",
  var_width = "4cm",
  desc_width = "11cm"
)
```

All other variables (from `01_uci_preprocessed.csv`) remain unchanged.

```{r, message = FALSE}
uci_random <- read_csv(file.path(work_dir, "shopping", "data", "intermediate_data",
                                 "02_uci_random_censor.csv"),
                       show_col_types = FALSE)
glimpse(uci_random)
```
